Sparse spectral methods for partial differential equations on spherical caps

### 2020-10-28 papers/spherical-caps/spherical-caps.tex
Edits to spherical cap paper. Added comments and new figs

 \begin{figure}[t]
	\centering % <-- added
	\includegraphics[scale=0.3]{complexity-Nend=1000}
	\caption{Complexity plot showing the length of time taken to construct the operator matrix for $\DeltaS + V(x,y,z)$, with rotationally invariant potential $V(x,y,z) = \cos z$. \sotodoinline{Move to experiments section} \sotodoinline{Add solve time}}
	\centering
	\label{fig:complexity}
\end{figure}

@@

\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:construction}
	Let $w : (\alpha,\beta) \: \to \R$ be a weight function. For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$. Then the 3D polynomials defined on $\Omega$
\begin{align*}
	\scopnki(x,y,z) \equiv \scopnki(\xvec,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr(\frac{\xvec}{\rho(z)})
\end{align*}
for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
\begin{align*}
	\ip< p, \: q > &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: w(z) \: \D \sigma(\xvec, z) \\
	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: w(z) \: \D z
\end{align*}
on $\Omega$, where $\D \sigma(\xvec, z)$ is the Lebesgue measure\sotodo{Is Lebesgue measure the right term? From my understanding the measure of a sphere would be 0} on $\Omega$. 
\end{proposition}

@@

The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Let $N \in \N$ and define notation that groups degree $N$ polynomials together[: 
\begin{align}
	\bigscopNka &:= 
		\begin{pmatrix}
			\scopa_{k,k,0}(x,y,z) \\
			\scopa_{k,k,1}(x,y,z) \\
			\vdots \\
			\scopa_{N,k,0}(x,y,z) \\
			\scopa_{N,k,1}(x,y,z)
		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
	\bigscopa_{N,0} &:= 
		\begin{pmatrix}
			\scopa_{0,0,0}(x,y,z) \\
			\vdots \\
			\scopa_{N,0,0}(x,y,z) \\
		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
	\bigscopNa &:= 
		\begin{pmatrix}
			\bigscopa_{N,0} \\
			\vdots \\
			\bigscopa_{N,N}
		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
\end{align}
and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to\sotodo{The following is wrong as it only applies if $N = \infty$}
\begin{align}
	J_x^{(a)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
	J_y^{(a)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
	J_z^{(a)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
\end{align}
The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis. For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated up to order $N \in \N$ by its expansion $f(x,y,z) = \bigscopNa(x,y,z)^\top \mathbf{f}$. Then $x \: f(x,y,z)$ is approximated by $\bigscopNa(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$. In other words, ${J_x^{(a)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis. Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:

@@

In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \enquote{Clenshaw matrices} -- we define the Clenshaw matrices as the matrices derived from the OP's $x,y,z$-multiplication relationships that form the multidimensional three-term recurrences. While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the \enquote{Clenshaw matrices} necessary for efficient function evaluation using the Clenshaw algorithm. In practice, it is simply a matter of converting coefficients between the two orderings.

Define, for $n=0,1,2,\dots$:  \sotodo{Explain why these are needed}
\begin{align*}
	\bigscoptna := 
		\begin{pmatrix}
			\scopa_{n,0,0}(x,y,z) \\
			\scopa_{n,1,0}(x,y,z) \\
			\scopa_{n,1,1}(x,y,z) \\
			\vdots \\
			\scopa_{n,n,0}(x,y,z) \\
			\scopa_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2n+1}, 
	\quad \quad 
	\bigscopta := 
		\begin{pmatrix}
			\bigscopta_0 \\
			\bigscopta_1 \\
			\bigscopta_2 \\
			\vdots \\
		\end{pmatrix}
\end{align*}
and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to \sotodo{Can the following be simplified? Seems like it repeats}
\begin{align}
	\tilde J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	\tilde J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
	\tilde J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
\end{align}
where
\begin{align*}
	\tilde J_{x/y/z}^{(a)} &= 
		\begin{pmatrix}
			\tilde B^{(a)}_{x/y/z, 0} & \tilde A^{(a)}_{x/y/z, 0} & & & & \\
			\tilde C^{(a)}_{x/y/z, 1} & \tilde B^{(a)}_{x/y/z, 1} & \tilde A^{(a)}_{x/y/z, 1} & & & \\
			& \tilde C^{(a)}_{x/y/z, 2} & \tilde B^{(a)}_{x/y/z, 2} & \tilde A^{(a)}_{x/y/z, 2} & & & \\
			& & \tilde C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
\end{align*}
These matrices are then block-tridiagonal (block-bandwidths $(1,1)$).

@@

Note that we can define an explicit \(\Dnt\) as follows:
\begin{align*}
	\Dnt := 
		\begin{pmatrix}
			0 & & & 0 & & & (\Gamma^{(a)}_{n,0,3})^{-1} & & \\
			& \ddots & & & \ddots & & & \ddots \\
			& & 0 & & & 0 & & & (\Gamma^{(a)}_{n,n,3})^{-1}  \\
			& & & & & \bm{\eta}^\top_{0} & & & \\
			& & & & & \bm{\eta}^\top_{1} & & &
		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
\end{align*}
for $n = 1, 2, \dots$ where again $\Gamma^{(a)}_{n,k,3}$ are defined in equations (\ref{eqn:jacobisubblocksGamma1}, \ref{eqn:jacobisubblocksGamma2}) for $k=0,\dots,n$, and where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
\begin{align*}
	\big(\bm{\eta}_{0}\big)_j &= 
		\begin{cases}
			\frac{1}{\betaa_{n,n,1,6}} & j = 2(2n+1) \\
			\frac{- \: \betaa_{n,n,1,5}}{\betaa_{n,n,1,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 3 \\
			0 & o/w
		\end{cases} \\
	\big(\bm{\eta}_{1}\big)_j &= 
		\begin{cases}
			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2  \text{ and } n > 1 \\
			0 & o/w
		\end{cases}
\end{align*}
For $n=0$, we can simply take
\begin{align*}
	D^\top_0 &:= 
		\begin{pmatrix}
			0 & 0 & \frac{1}{\gammaa_{0,0,3}} \\
			\frac{1}{\alphaa_{0,0,6}} & 0 & 0 \\
			0 & \frac{1}{\betaa_{0,0,6}} & 0
		\end{pmatrix} \in \R^{3\times3}.
\end{align*}

@@

\begin{definition}\label{def:clenshawmats}
The Clenshaw matrices associated with the OPs $\{\scopnkia\}$ are given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
\end{definition}

@@

\begin{theorem}\label{theorem:sparsityofdifferentialoperators}
	The operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure. More specifically:
\begin{itemize}
	\item $D_\varphi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 4)$
  	\item $W_\varphi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 2)$
	\item $D_\theta$ has  block-bandwidths $(0,0)$, and sub-block-bandwidths $(1, 1)$
	\item $\calL^{(a)\to(a+\tilde a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(0, 4)$
	\item $\calL_W^{(a)\to(a-\tilde a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 0)$
	\item $\Delta^{(1)}_W$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 2)$
\end{itemize}
\end{theorem}

@@

Finally, fix $a = 1$. For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are
\begin{align*}
	c_{m,j,h} = \norm{\scopmjh^{(1)}}^{-2}_{\scop^{(1)}} \ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}}.
\end{align*}
Using \bsrefeqn{eqn:woplaplacian} with $a = 1$, and \bsreflemma{lemma:Rsecondderivative}, we then have that
\begin{align*}
	&\ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}} \\
	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(1,2j)} \: \Big\{ \genjacnmk^{(a, 2k)} \: [- k^2 \genjacw^{(1, k)} - \genjacw^{(1, k)} - 2(k + 1)z \genjacw^{(0, k)}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a, 2k) \: \prime} \: [2 \genjacw^{(0, k+2)} - 2(k + 1)z \genjacw^{(1, k)}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a, 2k) \: \prime \prime} \: \genjacw^{(1, k+2)} \Big\} \: \genjacw^{(1, j)} \: \D z \Big) \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(1,2k)} \: \Big\{ \genjacnmk^{(a, 2k)} [-k(k+1) \genjacw^{(1, 0)} - 2(k+1)z + c_{n,k}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + c_{n-1,k} \genjac_{n-k-1}^{(a, 2k)} + c_{n+1,k} \genjac_{n-k+1}^{(a, 2k)} \Big\} \: \genjacw^{(1, 2k)} \: \D z \\
	&\quad = - \pi \: \delta_{k,j} \: \delta_{i,h} (\delta_{m,n-1} + \delta_{m,n} + \delta_{m,n+1}) \: \int_\alpha^1 \: \Big\{ \genjac_{m-k}^{(1,2k)} \: \genjacnmk^{(a, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\ 
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(a,2k) \: \prime} \: \genjac_{m-k}^{(a,2k) \: \prime} \: \genjacw^{(a+1, 2(k+1))} \Big\} \: \D z
\end{align*}
where the $c_{n-1,k}, c_{n,k}, c_{n+1,k}$ are those derived in \bsreflemma{lemma:Rsecondderivative}.
\end{proof}

There exist conversion matrix operators that increment/decrement the parameters, transforming the OPs from one (weighted or non-weighted) parameter space to another.

\begin{definition}\label{def:parametertransformationoperators}
Define the operator matrices $T^{(a)\to(a+\tilde a)}, \quad T_W^{(a)\to(a-\tilde a)}$ for conversion between non-weighted spaces and weighted spaces respectively according to
\begin{align*}
	\bigscopNa(x,y,z) &= \Big(T^{(a)\to(a+\tilde a)} \Big)^\top \: \bigscopN^{(a+\tilde a)}(x,y,z) \\
	\bigWNa(x,y,z) &= \Big(T_W^{(a)\to(a-\tilde a)} \Big)^\top \: \bigW_N^{(a-\tilde a)}(x,y,z)
\end{align*}
\end{definition}

@@

\begin{figure}[t]
	\centering % <-- added
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[scale=0.3]{sparsity-of-laplacian}
		\centering
		\caption{The Laplace operator $\Delta^{(1)}_W$}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[scale=0.3]{sparsity-of-rho2laplacian}
		\centering
		\caption{The $\rho^2$-factored Laplacian $L_W^{(1)}$}
	\end{subfigure}\hfil % <-- added

	\medskip
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[scale=0.3]{sparsity-of-helmholtz}
		\centering
		\caption{A variable coefficient Helmholtz operator}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[scale=0.23]{sparsity-of-biharmonic}
		\centering
		\caption{The Biharmonic operator $B_W^{(2)}$} 
	\end{subfigure}\hfil % <-- added
	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity. For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{(x_0^2 + z_0^2)}$ \sotodoinline{Shouldn't this be $\sqrt{1-x_0^2-y_0^2}$?} and $k = 200$.}
	\label{fig:sparsity}
\end{figure}

>>>>

@@

\begin{proposition}[\cite{olver2020orthogonal}]\label{prop:construction}
	Let $w : (\alpha,\beta) \: \to \R$ be a weight function. For $n = 0,1,2,\dots, $ let $\{r_{n,k}\}$ be polynomials orthogonal with respect to the weight $\rho(x)^{2k} w(x)$ where $0 \le k \le n$. Then the 3D polynomials defined on $\Omega$
\begin{align*}
	\scopnki(x,y,z) \equiv \scopnki(\xvec,z) := r_{n-k,k}(z) \: \rho(z)^k \: \ch_{k,i}\fpr(\frac{\xvec}{\rho(z)})
\end{align*}
for $i \in {0,1}, \: 0 \le k \le n, \: n = 0,1,2,\dots$ are orthogonal polynomials with respect to the inner product
\begin{align*}
	\ip< p, \: q > &:= \int_\Omega p(x,y,z) \: q(x,y,z) \: w(z) \: \D A \\
	&= \int_0^{\cos^{-1}(\alpha)} \int_0^{2\pi} p\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: q\big(\sinphi \costheta, \sinphi \sintheta, \cosphi \big) \: w(\cosphi) \: \sinphi \: \D \theta \: \D \varphi \\
	&= \int_\alpha^1 \int_0^{2\pi} p\big(\rho(z) \cos\theta, \rho(z) \sin\theta, z\big) \: q\big(\rho(z) \cos\theta, \rho(z) \sin \theta, z\big) \: w(z) \: \D \theta \: \D z
\end{align*}
on $\Omega$, where $\D A = \sinphi \: \D \theta \: \D \varphi$ is the Lebesgue measure on $\Omega$. 
\sotodoinline{Is Lebesgue measure the right term? From my understanding the measure of a sphere would be 0} 
\bstodoinline{I have changed it to dA and elaborated a bit more on it, but I am not sure on this so you could be correct that another term needs to be used. Your quadratic surfaces paper with Yuan says Lebesgue I think when discussing this. I have tried to keep it general here by using a measure definition, but we could omit this if its just going to cause confusion?}
\end{proposition}

@@

The recurrences in Lemma \ref{lemma:Qrecurrence} lead to Jacobi operators that correspond to multiplication by $x$, $y$ and $z$. Let $N \in \N$ and define notation that groups degree $N$ polynomials together[: 
\begin{align}
	\bigscopNka &:= 
		\begin{pmatrix}
			\scopa_{k,k,0}(x,y,z) \\
			\scopa_{k,k,1}(x,y,z) \\
			\vdots \\
			\scopa_{N,k,0}(x,y,z) \\
			\scopa_{N,k,1}(x,y,z)
		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:OPdefNka} \\ 
	\bigscopa_{N,0} &:= 
		\begin{pmatrix}
			\scopa_{0,0,0}(x,y,z) \\
			\vdots \\
			\scopa_{N,0,0}(x,y,z) \\
		\end{pmatrix} \in \R^{N+1}, \label{eqn:OPdefN0a} \\
	\bigscopNa &:= 
		\begin{pmatrix}
			\bigscopa_{N,0} \\
			\vdots \\
			\bigscopa_{N,N}
		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:OPdefNa}
\end{align}
and set $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ as the Jacobi matrices corresponding to
\sotodoinline{The following is wrong as it only applies if $N = \infty$} 
\bstodoinline{What is better? To define them in this way for $N \to \infty$? Or to include the extra terms in this defn below? Or just state that the Jacobi operators approximate multiplication by x,y,z? This also relates to the next comments}
\begin{align}
	J_x^{(a)} \: \bigscopNa(x,y,z) = x \: \bigscopNa(x,y,z), \nonumber \\
	J_y^{(a)} \: \bigscopNa(x,y,z) = y \: \bigscopNa(x,y,z), \label{eqn:jacobimatricesdefinition} \\
	J_z^{(a)} \: \bigscopNa(x,y,z) = z \: \bigscopNa(x,y,z). \nonumber
\end{align}
The matrices $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ act on the coefficients vector of a function's expansion in the $\{\scopnkia\}$ basis. For example, let $a$ be a general parameter and let a function $f(x,y,z)$ defined on $\Omega$ be approximated up to order $N \in \N$ by its expansion $f(x,y,z) = \bigscopNa(x,y,z)^\top \mathbf{f}$. Then $x \: f(x,y,z)$ is approximated by $\bigscopNa(x,y,z)^\top {J_x^{(a)\top}} \mathbf{f}$. In other words, ${J_x^{(a)\top}} \mathbf{f}$ is the coefficients vector for the expansion of the function $(x,y,z) \mapsto x \: f(x,y,z)$ in the  $\{\scopnkia\}$ basis. Note that $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are banded-block-banded matrices:

@@

In this section, we introduce new notation for a different ordering of the OP vector, in order to demonstrate how we can obtain a recurrence for the OPs in matrix-vector form, and derive the associated \textit{recurrence coefficient matrices} -- we define these as the matrices derived from the OP's $x,y,z$-multiplication relationships that form the coefficients for the multidimensional three-term recurrences.

While in Section \ref{subsection:jacobimats} the ordering used is convenient for the application of Jacobi and differential operators to the vector of coefficients of a given function's expansion, the ordering we will use in this section is convenient for building the OPs and more importantly obtaining the recurrence coefficient matrices necessary for efficient function evaluation using the Clenshaw algorithm. In practice, it is simply a matter of converting coefficients between the two orderings. To this end, we define our OP-building ordering as follows. For $n=0,1,2,\dots$:
\begin{align*}
	\bigscoptna := 
		\begin{pmatrix}
			\scopa_{n,0,0}(x,y,z) \\
			\scopa_{n,1,0}(x,y,z) \\
			\scopa_{n,1,1}(x,y,z) \\
			\vdots \\
			\scopa_{n,n,0}(x,y,z) \\
			\scopa_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2n+1}, 
	\quad \quad 
	\bigscopta := 
		\begin{pmatrix}
			\bigscopta_0 \\
			\bigscopta_1 \\
			\bigscopta_2 \\
			\vdots \\
		\end{pmatrix}
\end{align*}
and set $\tilde J_x^{(a)}, \tilde J_y^{(a)}, \tilde J_z^{(a)}$ as the Jacobi matrices corresponding to 
\sotodoinline{Can the following be simplified? Seems like it repeats} 
\bstodoinline{Is it worth removing the previous? And just explaining that we can reorder the Jacobi matrices according to the OP/coefficient vector ordering? This section is important for deriving how to build the OPs, whereas before I simply need to say that the Jacobi operators that apply to coefficients vectors are reorderings of the ones in this section.}
\begin{align}
	\tilde J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	\tilde J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:jacobimatricesdefinitionalt} \\
	\tilde J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
\end{align}
where
\begin{align*}
	\tilde J_{x/y/z}^{(a)} &= 
		\begin{pmatrix}
			\tilde B^{(a)}_{x/y/z, 0} & \tilde A^{(a)}_{x/y/z, 0} & & & & \\
			\tilde C^{(a)}_{x/y/z, 1} & \tilde B^{(a)}_{x/y/z, 1} & \tilde A^{(a)}_{x/y/z, 1} & & & \\
			& \tilde C^{(a)}_{x/y/z, 2} & \tilde B^{(a)}_{x/y/z, 2} & \tilde A^{(a)}_{x/y/z, 2} & & & \\
			& & \tilde C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
\end{align*}
These matrices are then block-tridiagonal (block-bandwidths $(1,1)$).

@@

Note that we can define an explicit $\Dnt$ as follows:
\begin{align*}
	\Dnt := 
		\begin{pmatrix}
			0 & & & 0 & & & (\Gamma^{(a)}_{n,0,3})^{-1} & & \\
			& \ddots & & & \ddots & & & \ddots \\
			& & 0 & & & 0 & & & (\Gamma^{(a)}_{n,n,3})^{-1}  \\
			& & & & & \bm{\eta}^\top_{0} & & & \\
			& & & & & \bm{\eta}^\top_{1} & & &
		\end{pmatrix} \in \R^{(2n+3)\times3(2n+1)},
\end{align*}
for $n = 1, 2, \dots$ where again $\Gamma^{(a)}_{n,k,3}$ are defined in equations (\ref{eqn:jacobisubblocksGamma1}, \ref{eqn:jacobisubblocksGamma2}) for $k=0,\dots,n$, and where $\bm{\eta}_{0}, \bm{\eta}_{1} \in \R^{3(2n+1)}$ with entries given by 
\begin{align*}
	\big(\bm{\eta}_{0}\big)_j &= 
		\begin{cases}
			\frac{1}{\betaa_{n,n,1,6}} & j = 2(2n+1) \\
			\frac{- \: \betaa_{n,n,1,5}}{\betaa_{n,n,1,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 3 \\
			0 & o/w
		\end{cases} \\
	\big(\bm{\eta}_{1}\big)_j &= 
		\begin{cases}
			\frac{1}{\alphaa_{n,n,6}} & j = 2n+1 \\
			\frac{- \: \alphaa_{n,n,5}}{\alphaa_{n,n,6} \: \gammaa_{n, n-1, 3}} & j = 3(2n+1) - 2  \text{ and } n > 1 \\
			0 & o/w
		\end{cases}
\end{align*}
For $n=0$, we can simply take
\begin{align*}
	D^\top_0 &:= 
		\begin{pmatrix}
			0 & 0 & \frac{1}{\gammaa_{0,0,3}} \\
			\frac{1}{\alphaa_{0,0,6}} & 0 & 0 \\
			0 & \frac{1}{\betaa_{0,0,6}} & 0
		\end{pmatrix} \in \R^{3\times3}.
\end{align*}

@@

\begin{definition}\label{def:clenshawmats}
	The \textit{recurrence coefficient matrices} associated with the OPs $\{\scopnkia\}$ are given by the matrices $A_n, B_n, C_n, \Dnt$ for $n = 0,1,2,\dots$ defined above.
\end{definition}

@@

\begin{theorem}\label{theorem:sparsityofdifferentialoperators}
	The operator matrices $D_\varphi^{(a)}, \: W_\varphi^{(a)}, \: D_\theta, \: \calL^{(a)\to(a+\tilde a)}, \: \calL_W^{(a)\to(a-\tilde a)}, \: \Delta^{(1)}_W$ from Definition \ref{def:differentialoperators} are sparse, with banded-block-banded structure. More specifically:
\begin{itemize}
	\item $D_\varphi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 4)$
  	\item $W_\varphi^{(a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 2)$
	\item $D_\theta$ has  block-bandwidths $(0,0)$, and sub-block-bandwidths $(1, 1)$
	\item $\calL^{(a)\to(a+\tilde a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(0, 4)$
	\item $\calL_W^{(a)\to(a-\tilde a)}$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(4, 0)$
	\item $\Delta^{(1)}_W$ has block-bandwidths $(0,0)$, and sub-block-bandwidths $(2, 2)$
\end{itemize}
\end{theorem}

@@

Finally, fix $a = 1$. For the operator $\Delta^{(1)}_W$ for the Laplacian on the weighted space, the entries of the operator are given by the coefficients in the expansion $\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big) = \sum_{m=0}^{n+2} \sum_{j=0}^m \sum_{h=0}^1 c_{m,j,h} \: \scopmjh^{(1)}$, where the coefficients are given by
\begin{align*}
	c_{m,j,h} = \norm{\scopmjh^{(1)}}^{-2}_{\scop^{(1)}} \ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}}.
\end{align*}
Using \bsrefeqn{eqn:woplaplacian} with $a = 1$, and \bsreflemma{lemma:Rsecondderivative}, we then have that
\begin{align*}
	&\ip<\DeltaS \big(\genjacw^{(1,0)} \: \scopnki^{(1)} \big), \: \scopmjh^{(1)}>_{\scop^{(1)}} \\
	&\quad = \Big( \int_0^{2\pi} \: \chki(\theta) \: \chjh(\theta) \: \D \theta \Big) \\
	&\quad \quad \quad \quad \cdot \: \Big( \int_\alpha^1 \: \genjacmmj^{(1,2j)} \: \Big\{ \genjacnmk^{(1, 2k)} \: [- k^2 \genjacw^{(1, k)} - \genjacw^{(1, k)} - 2(k + 1)z \genjacw^{(0, k)}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1, 2k) \: \prime} \: [2 \genjacw^{(0, k+2)} - 2(k + 1)z \genjacw^{(1, k)}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1, 2k) \: \prime \prime} \: \genjacw^{(1, k+2)} \Big\} \: \genjacw^{(1, j)} \: \D z \Big) \\
	&\quad = \pi \: \delta_{k,j} \: \delta_{i,h} \: \int_\alpha^1 \: \genjac_{m-k}^{(1,2k)} \: \Big\{ \genjacnmk^{(1, 2k)} [-k(k+1) \genjacw^{(1, 0)} - 2(k+1)z + c_{n,k}] \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + c_{n-1,k} \genjac_{n-k-1}^{(1, 2k)} + c_{n+1,k} \genjac_{n-k+1}^{(1, 2k)} \Big\} \: \genjacw^{(1, 2k)} \: \D z \\
	&\quad = - \pi \: \delta_{k,j} \: \delta_{i,h} (\delta_{m,n-1} + \delta_{m,n} + \delta_{m,n+1}) \: \int_\alpha^1 \: \Big\{ \genjacnmk^{(1,2k)} \: \genjac_{m-k}^{(1, 2k)} (k(k+1) \genjacw^{(1, 0)} + 2(k+1)z) \\ 
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad \quad + \genjacnmk^{(1,2k) \: \prime} \: \genjac_{m-k}^{(1,2k) \: \prime} \: \genjacw^{(2, 2(k+1))} \Big\} \: \D z
\end{align*}
where the $c_{n-1,k}, c_{n,k}, c_{n+1,k}$ are those derived in \bsreflemma{lemma:Rsecondderivative}.
\end{proof}

By applying these differential operators, we are (in some cases) incrementing or decrementing the parameter value $a$. It is therefore necessary to also be able to raise or lower the parameter by way of an independent operator. There exist conversion matrix operators that do exactly this, transforming the OPs from one (weighted or non-weighted) parameter space to another.
\begin{definition}\label{def:parametertransformationoperators}
Define the operator matrices $T^{(a)\to(a+\tilde a)}, \quad T_W^{(a)\to(a-\tilde a)}$ for conversion between non-weighted spaces and weighted spaces respectively according to
\begin{align*}
	\bigscopNa(x,y,z) &= \Big(T^{(a)\to(a+\tilde a)} \Big)^\top \: \bigscopN^{(a+\tilde a)}(x,y,z) \\
	\bigWNa(x,y,z) &= \Big(T_W^{(a)\to(a-\tilde a)} \Big)^\top \: \bigW_N^{(a-\tilde a)}(x,y,z)
\end{align*}
\end{definition}

@@

\begin{figure}[t]
	\centering % <-- added
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[scale=0.3]{sparsity-of-laplacian}
		\centering
		\caption{The Laplace-Beltrami operator $\Delta^{(1)}_W$}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[scale=0.3]{sparsity-of-rho2laplacian}
		\centering
		\caption{The $\rho^2$-factored Laplace-Beltrami operator $D_\varphi^{(0)} \: W_\varphi^{(1)} + T^{(0)\to(1)} \: T_W^{(1)\to(0)} \: (D_\theta)^2$}
	\end{subfigure}\hfil % <-- added

	\medskip
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[scale=0.3]{sparsity-of-helmholtz}
		\centering
		\caption{A variable coefficient Helmholtz operator}
	\end{subfigure}\hfil % <-- added
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[scale=0.23]{sparsity-of-biharmonic}
		\centering
		\caption{The Biharmonic operator $\calB_W^{(2)}$} 
	\end{subfigure}\hfil % <-- added
	\caption{\enquote{Spy} plots of (differential) operator matrices, showing their sparsity. For (c), the the weighted variable coefficient Helmholtz operator is $\Delta^{(1)}_W + k^2 \: T^{(0)\to(1)} \: V({J_x^{(0)}}^\top, {J_y^{(0)}}^\top, {J_z^{(0)}}^\top) \: T_W^{(1)\to(0)}$ for $v(x,y,z) = 1 - (3(x-x_0)^2 + 5(y-y_0)^2 + 2(z-z_0)^2)$ where $(x_0, z_0) := (0.7, 0.2)$, $y_0 := \sqrt{1 - x_0^2 - z_0^2}$ and $k = 200$.}
	\label{fig:sparsity}
\end{figure}
